Deterministic Scheduling of Periodic Messages by an SDN Approach

### 2017-07-20 infocom/infocom.tex
temps 16 routes
figures vspace
trois modifs en relisant

      In the process we study, a message is sent on each route at each period, denoted by $P$.
      Let $r$ be a route, if a message is sent at time $m$ from $s$ the first vertex of $r$ then it will arrive at vertex $v$ in $r$ at time $m + \lambda(v,r)$. Since the process is periodic, if we look at the messages going through an arc between time $0$ and $P-1$, then we would observe the same messages in the same order and position between time $kP$ and $(k+1)P -1$ for all $k$. 
      Therefore, the first time slot at which a message sent at time $m$ on $r$ reaches a vertex $v$ in $r$ is $t(v,r) = m + \lambda(v,r) \mod P$.

@@

Finally, there was no collisions between routes at the beginning and since we modify $\mathcal{M}$ only if it creates no collision, the assignment we obtain at the end has no collisions between routes.
\end{proof}

@@

\begin{theorem}\label{th:FPT}
$PALL \in \FPT$ when parametrized by the number of routes.
\end{theorem}
\begin{proof}
Let $(G, {\cal R})$ be a star routed network and let $\mathcal{M}$ be a $(P,\tau)$-periodic assignment of $(G, {\cal R})$. First remark that for a given assignment and a route $r_0$ with offset $m$, by removing $m$ to all offsets, we can always assume that its offset is zero. Therefore we need only to consider all \emph{compact assignments} with an \emph{offset $0$} for the route $r_0$. 
We now evaluate the number of compact assignments and prove that it only depends
on $n$ the number of routes which proves the theorem. 
We give a way to build a compact assignment $\mathcal{M}$ by determining its offsets one after the other. We fix an arbitrary total order on ${\cal R})$.
First a route $r_0$ is chosen arbitrarily and its offset set to $0$. 
Then at each step, if the offsets of $S \subseteq  {\cal R}$ have been chosen,
we select the smallest route $r$ in $S$ for the order. 
Then we select a route in $r' \in {\cal R} \setminus S$ and set its offset such that 
if we remove $1$ then $r'$ collides with $r$. Note that if $r$ is a forward route (resp. a backward route) then $r'$ is also a forward route (resp. a backward route). We can also decide to 
not use route $r$. At a given step of the algorithm, if $|S| = 2i$, we have $n-i$ choices 
of routes to select and the value of the offset is determined by the values of the offsets of routes in $S$. Therefore there at most $n!$ different compact assignments with offset $r_0$
fixed to $0$.

An algorithm to solve PALL build every possible compact assignment as described here, and
test at each step whether there is a collision which can be done in time linear in the size of 
$(G, {\cal R})$. Therefore $PALL \in \FPT$.
\end{proof}

We call the algorithm described in Theorem~\ref{th:FPT} \textbf{Exhaustive Search of Compact Assignments (ESCA)}. To make it more efficient in practice, we make cuts in the search tree 
used to explore all compact assignments. Consider a set of $k$ forward routes whose offsets has been fixed at some point in the search tree. We consider the times at which the messages of these routes cross the central arc. It divides the period into $[(a_0,b_0), \dots, (a_{k-1},b_{k-1})]$ such that the central arc is free only during the intervals $(a_i,b_i)$. 
Therefore at most $\displaystyle{ \sum_{i=0}^{k} \lfloor(b_{i} -a_i)/\tau\rfloor} $ forward routes can still use the central arc. If this value is more than $n - k$, it is not possible to create a compact assignment by extending the one on $S$ and we backtrack in the search tree. We do the same cut for the backward routes.

@@

@@

      \begin{center}
	 \hspace{-0.5cm}\includegraphics[width=0.53\textwidth]{periode_petite.pdf}
      \end{center}
      \caption{Minimum period averaged over $1000$ random instances}
      \end{figure}

@@

       \begin{center}
      \hspace{-0.5cm}\includegraphics[width=0.5\textwidth]{echec_longues.pdf}
      \end{center}
      \caption{Success rate for $8$ routes over $1000$ random instances}
     \end{figure}

@@

      The computation time of the exhaustive search is in $O(n!)$, hence we need to study its scalability when $n$ grows. In the following experiment, the weights of the arcs $(c_t,t_i)$ are drawn following an uniform distribution between $0$ and $30000$ slots. We fixed a period of $1.05n\tau$, which corresponds to $95\%$ of load.
%       , that corresponds to loads of $\simeq 70\%$ (for $1$ route) to $\simeq 98\%$ (for 21 routes).
      The following table shows the time before the exhaustive search ends, for $8$ to $16$ routes, averaged on $100$ random star routed networks. This shows that for less than $15$ routes, which corresponds to all current topologies, the algorithm is efficient enough, but on more routes we should improve it further.
         \begin{center}
         \begin{tabularx}{0.5\textwidth}{|l|X|X|X|X|}
    \hline
   $n$ & $8$ & $10$& $12$& $16$\\
    \hline
   Time (s) & $6.10^{-5}$&$8.10^{-4}$&$2.10^{-2}$& $15$\\
    \hline
      \end{tabularx}
      \end{center}
         \section{Solving PALL on Star Routed Networks}\label{sec:PALL}

@@

     \begin{algorithm}[H]
    \caption{ Greedy deadline ({\bf GD}) }
    \begin{algorithmic}
    \REQUIRE A routed network $(G,{\cal R})$, a period $P$, packet size $\tau$, $ T_{max}$, the offsets $m_i$
    \ENSURE $(P,\tau)$-periodic assignment of $(G,{\cal R})$, or failure
   \STATE  ${\cal H} \leftarrow$ empty set //{\em set of eligible routes}
       \STATE  free\_ intervals $\leftarrow$ [0,$P$] //{\em list of intervals of free slots}

@@

  \centering
     \hspace{-1.5cm} \includegraphics[width=0.4\textwidth]{departs_gp_21000.pdf}
\vspace{-0.75cm}
      \caption{Success rate of different orders, $95\%$ load.}
          \end{figure}
          \vspace{-1cm}
\begin{figure}[H] 
          \includegraphics[width=0.5\textwidth]{departs_gp_25000.pdf}
          \vspace{-1cm}
      \caption{Success rate of different orders, $80\%$ load.}
           \label{fig:success30000}
     \end{figure}

@@

    \begin{figure} [H] 
       \begin{center}
      \includegraphics[width=0.5\textwidth]{retour_21000.pdf}
      \end{center}
      \caption{Success rate of GD, MLS and PMLS for $P = 21000$}
     \label{fig:success21000}
     \end{figure}

@@

       \begin{center}
      \includegraphics[width = 0.5\textwidth]{stochastic.pdf}
       %\input{figures/stochastic}
      \end{center}
      \caption{Needed margin with respect to the period for a stochastic policy}
      \label{fig:sto}
     \end{figure}

>>>>

      In the process we study, a message is sent on each route at each period, denoted by $P$.
      Let $r$ be a route, if a message is sent at time $m$ from $s$ the first vertex of $r$ then it will arrive at vertex $v$ in $r$ at time $m + \lambda(v,r)$. Since the process is periodic, if we look at the messages going through an arc between time $0$ and $P-1$, then we see the same messages in the same order and position between time $kP$ and $(k+1)P -1$ for all $k$. 
      Therefore, the first time slot at which a message sent at time $m$ on $r$ reaches a vertex $v$ in $r$ is $t(v,r) = m + \lambda(v,r) \mod P$.

@@

Finally, there is no collisions between routes at the beginning and since we modify $\mathcal{M}$ only if it creates no collision, the assignment we obtain at the end has no collisions between routes.
\end{proof}

@@

\begin{theorem}\label{th:FPT}
$PALL \in \FPT$ when parametrized by the number of routes.
\end{theorem}
\begin{proof}
Let $(G, {\cal R})$ be a star routed network and let $\mathcal{M}$ be a $(P,\tau)$-periodic assignment of $(G, {\cal R})$. First remark that for a given assignment and a route $r_0$ with offset $m$, by removing $m$ to all offsets, we can always assume that its offset is zero. Therefore we need only to consider all \emph{compact assignments} with an \emph{offset $0$} for the route $r_0$. 
We now evaluate the number of compact assignments and prove that it only depends
on $n$ the number of routes which proves the theorem. 
We give a way to build a compact assignment $\mathcal{M}$ by determining its offsets one after the other. We fix an arbitrary total order on ${\cal R}$.
First a route $r_0$ is chosen arbitrarily and its offset set to $0$. 
Then at each step, if the offsets of $S \subseteq  {\cal R}$ have been chosen,
we select the smallest route $r$ in $S$ for the order. 
Then we select a route in $r' \in {\cal R} \setminus S$ and set its offset such that 
if we remove $1$ then $r'$ collides with $r$. Note that if $r$ is a forward route (resp. a backward route) then $r'$ is also a forward route (resp. a backward route). We can also decide to 
not use route $r$. At a given step of the algorithm, if $|S| = 2i$, we have $n-i$ choices 
of routes to select. The value of the offset of the selected route is entirely determined by the values of the offsets of routes in $S$. Therefore there are at most $n!$ different compact assignments with offset $r_0$ fixed to $0$.

The algorithm to solve PALL builds every possible compact assignment as described here, and
tests at each step whether there is a collision which can be done in time linear in the size of 
$(G, {\cal R})$. Therefore $PALL \in \FPT$.
\end{proof}

We call the algorithm described in Theorem~\ref{th:FPT} \textbf{Exhaustive Search of Compact Assignments}. To make it more efficient in practice, we make cuts in the search tree used to explore all compact assignments. Consider a set of $k$ forward routes whose offsets has been fixed at some point in the search tree. We consider the times at which the messages of these routes cross the central arc. It divides the period into $[(a_0,b_0), \dots, (a_{k-1},b_{k-1})]$ such that the central arc is free only during the intervals $(a_i,b_i)$. Therefore at most $\displaystyle{ \sum_{i=0}^{k-1} \lfloor(b_{i} -a_i)/\tau\rfloor} $ forward routes can still use the central arc. If this value is more than $n - k$, it is not possible to create a compact assignment by extending the one on $S$ and we backtrack in the search tree. We do the same cut for the backward routes.

@@

      \todo{présenter ça comme une average load ?}

@@

      \begin{center}
	 \hspace{-0.5cm}\includegraphics[width=0.53\textwidth]{periode_petite.pdf}
      \end{center}
            \vspace{-1cm}
      \caption{Minimum period averaged over $1000$ random instances}
      \end{figure}

@@

       \begin{center}
      \hspace{-0.5cm}\includegraphics[width=0.5\textwidth]{echec_longues.pdf}
      \end{center}
            \vspace{-1cm}
      \caption{Success rate for $8$ routes over $1000$ random instances}
     \end{figure}

@@

      The computation time of the exhaustive search is in $O(n!)$, hence we need to study its scalability when $n$ grows. In the following experiment, the weights of the arcs $(c_t,t_i)$ are drawn following an uniform distribution between $0$ and $30000$ slots. We fixed a period of $1.05n\tau$, which corresponds to $95\%$ of load.
%       , that corresponds to loads of $\simeq 70\%$ (for $1$ route) to $\simeq 98\%$ (for 21 routes).
      The following table shows the time before the exhaustive search ends, for $8$ to $16$ routes, averaged on $100$ random star routed networks. This shows that for less than $15$ routes, which corresponds to all current topologies, the algorithm is efficient enough, but on more routes we should improve it further.
         \begin{center}
         \begin{tabularx}{0.5\textwidth}{|l|X|X|X|X|}
    \hline
   $n$ & $8$ & $10$& $12$& $16$\\
    \hline
   Time (s) & $6.10^{-5}$&$8.10^{-4}$&$2.10^{-2}$& $11$\\
    \hline
      \end{tabularx}
      \end{center}

         \section{Solving PALL on Star Routed Networks}\label{sec:PALL}

@@

     \begin{algorithm}
    \caption{ Greedy deadline ({\bf GD}) }
    \begin{algorithmic}
    \REQUIRE A routed network $(G,{\cal R})$, a period $P$, packet size $\tau$, $ T_{max}$, the offsets $m_i$
    \ENSURE $(P,\tau)$-periodic assignment of $(G,{\cal R})$, or failure
   \STATE  ${\cal H} \leftarrow$ empty set //{\em set of eligible routes}
       \STATE  free\_ intervals $\leftarrow$ [0,$P$] //{\em list of intervals of free slots}

@@

  \centering
     \hspace{-1.5cm} \includegraphics[width=0.4\textwidth]{departs_gp_21000.pdf}
\vspace{-0.75cm}
      \caption{Success rate of different orders, $95\%$ load.}
          \end{figure}
          \vspace{-1cm}
\begin{figure}[H] 
          \includegraphics[width=0.5\textwidth]{departs_gp_25000.pdf}
          \vspace{-1.5cm}
      \caption{Success rate of different orders, $80\%$ load.}
           \label{fig:success30000}
     \end{figure}

@@

    \begin{figure} [H] 
       \begin{center}
      \includegraphics[width=0.5\textwidth]{retour_21000.pdf}
      \end{center}
            \vspace{-1cm}
      \caption{Success rate of GD, MLS and PMLS for $P = 21000$}
     \label{fig:success21000}
     \end{figure}

@@

       \begin{center}
     \hspace{-0.5cm} \includegraphics[width = 0.5\textwidth]{stochastic.pdf}
       %\input{figures/stochastic}
      \end{center}
      \vspace{-1cm}
      \caption{Needed margin with respect to the period for a stochastic policy}
      \label{fig:sto}
     \end{figure}
